---
title: "Reproducible Research: Peer Assessment 1"
author: "Alberto Vicens Sanchez"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
knitr::opts_knit$set(root.dir = "~/cursos/data_science_Coursera/5_reproducible_research/RepData_PeerAssessment1/")
```

In this report, I describe the steps to analyze a dataset including the physical activity, given in number of steps taken in 5 minute intervals each day, of an anonymous individual during two months (October-November 2012). Through different steps, I will answer to different questions using the information provided in the dataset. 

## Loading and preprocessing the data
I downloaded the dataset from the Course web: 
[Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

I loaded the dataset, which is stored in *comma-separated-value (CSV)* format:

```{r load_data}
activity<-read.csv("activity.csv",header=TRUE, quote = "\"")
head(activity)

```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
