---
title: "Reproducible Research: Peer Assessment 1"
author: "Alberto Vicens Sanchez"
output: html_document
keep_md: true
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, cache=TRUE,cache.path="cache/", fig.path = "cache/", autodep=TRUE)
knitr::opts_knit$set(root.dir = "~/cursos/data_science_Coursera/5_reproducible_research/RepData_PeerAssessment1/")
```

In this report, I describe the steps to analyze a dataset including the physical activity, given in number of steps taken in 5 minute intervals each day, of an anonymous individual during two months (October-November 2012). Through different steps, I will answer to different questions using the information provided in the dataset. 

## Loading and preprocessing the data
I downloaded the dataset from the Course web: 
[Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

I load the dataset, which is stored in *comma-separated-value (CSV)* format:
```{r load_data}
activity<-read.csv("activity.csv",header=TRUE, quote = "\"")
str(activity)
```
I convert the date variable from factor to a date format
```{r convert_date}
activity$date<-as.Date(activity$date, format="%Y-%m-%d")
str(activity$date)
```

I also convert the interval from integer to time (hours: minutes)

```{r convert_time}
int2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(activity$interval))
int <- paste0(int2,activity$interval)
activity$interval<- format(strptime(int, format= "%H%M"), format ="%H:%M")
```

## What is mean total number of steps taken per day?

```{r totalStepsPerDay, fig.height=5, fig.width=8}
totalSteps<-tapply(activity$steps, activity$date, sum, na.rm=TRUE)
plot(totalSteps, type = "h", col="darkblue", xaxt="n", ylab="Total number of steps", lwd=2, xlab="", main="Total number of steps per day")
axis(1,at=1:61,labels=names(totalSteps), las=2, cex.axis=0.6)
```
```{r meanSteps}
meanSteps<-round(mean(totalSteps),digits = 2)
medianSteps<-median(totalSteps)
```
The mean steps per day was **`r meanSteps`** and the median **`r medianSteps`** 

## What is the average daily activity pattern?

```{r, averageDailySteps, echo=FALSE, fig.height=4, fig.width=8}
averageSteps<-tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(averageSteps, type = "l", col="darkblue", xaxt="n", ylab="Average number of steps", lwd=2, xlab="Time", main="Average number of steps per interval")
axis(1,at=c(1,73,145,217,288),labels=names(averageSteps[c(1,73,145,217,288)]), cex.axis=0.6)
```

```{r maxStepsInterval}
maxInterval<-names(averageSteps[averageSteps == max(averageSteps)])
```
The time with highest activity is `r maxInterval`

## Imputing missing values

```{r NAvalues}
naCount<-length(which(is.na(activity$steps)))
``` 
In the dataset there is a total of *`r naCount` NA values*.

To fill in the NA value, I will replace each NA value for the mean of that 5-day interval, creating the new variable *steps2*
```{r, replaceNAvalues}
for (i in 1:nrow(activity)) {
if (is.na(activity$steps[i])){
int<-activity[i,3]
activity$steps2[i] <- averageSteps[as.character(int)]
}
else { activity$steps2[i] <- activity$steps[i]}
}
head(activity)
```

```{r totalStepsPerDay2, fig.height=5, fig.width=8}
totalSteps2<-tapply(activity$steps2, activity$date, sum)
steps<-rbind(totalSteps,totalSteps2)
barplot(steps,beside = TRUE, col =c("blue","red"), xlab="", ylab="Total number of steps",main="Total number of steps per day", cex.axis=0.8, cex.names=0.6, las=2)
legend("top",c("with NA values","imputed NA values"), fill = c("blue","red"), cex = 0.8,bty = "n")
```
```{r meanSteps2}
meanSteps2<-round(mean(totalSteps2),digits = 2)
medianSteps2<-median(totalSteps2)
stepsMatrix<-matrix(c(meanSteps, meanSteps2,medianSteps,medianSteps2),nrow=2,ncol=2,dimnames=list(c("means","medians"),c("with NA","imputted NA")))
library(knitr)
kable(stepsMatrix)
```

There is difference in the mean but not in the of total number steps between the datasets.

## Are there differences in activity patterns between weekdays and weekends?
```{r classify_days}
activity$day<-weekdays(activity$date)
weekday<-c("lunes","martes","miércoles","jueves","viernes")

daytype = character()
for (j in 1:nrow(activity)) {
  if (activity$day[j] %in% weekday) {
    daytype[j] =  "weekday" 
    } else { daytype[j] = "weekend"}
  }
activity<-cbind(activity,daytype)
```

```{r compare_weekdays, fig.height=5, fig.width=8}
averageStepsbyDay<-tapply(activity$steps2, list(activity$interval, activity$daytype), mean, na.rm=T)

library(reshape)
averageStepsMelted<-melt.array(averageStepsbyDay, varnames=c("interval","daytype"))

library(ggplot2)
ggplot(data=averageStepsMelted,aes(x=interval,y=value, group=daytype)) + geom_line(aes(color = daytype), size = 1.2) + theme(legend.position = "top") + xlab("Time") + ylab("Number of steps") + scale_x_discrete(breaks=c("00:00","06:00","12:00","18:00","23:55")) + labs(title="Number of steps beetwen weekdays and weekend", color ="Day type")
```

